### 2022-01-31 ARSimulation.Rmd
Added Plot of Bootstrap

---
title: "Block Bootstrap for Time Series"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Simulates, prints, and plots auto-regressive process of order 1 with ar = 0.5 and size = 100
```{r, echo = FALSE}
ts.sim <- arima.sim(list(order = c(1, 0, 0), ar = 0.5), n = 100) 
ts.sim
plot(ts.sim)
```

Splits time series into ten blocks of size 10 and stores them in a list
```{r, echo = False}
ts.blocks <- list(ts.sim[c(1 : 10)], ts.sim[c(11 : 20)], ts.sim[c(21 : 30)], ts.sim[c(31 : 40)], ts.sim[c(41 : 50)], ts.sim[c(51 : 60)], ts.sim[c(61 : 70)], ts.sim[c(71 : 80)], ts.sim[c(81 : 90)], ts.sim[c(91 : 100)])
ts.blocks
```

Randomly samples 10 non-overlapping blocks (with replacement) from ts.blocks, creating a single bootstrapped time series
```{r, echo = False}
ts.boot1 <- sample(ts.blocks, 10, replace = T)
ts.boot1
class(ts.boot1)
```

ts.boot1 is a list of vectors and does not work with the plot() function, so we use a nested loop to store the values in ts.boot2, a simpler numeric vector
```{r, echo = False}
ts.boot2 <- c()
for(i in 1 : 10) {
  for(j in 1 : 10) {
    ts.boot2 <- c(ts.boot2, ts.boot1[[i]][[j]])
  }
}
ts.boot2
class(ts.boot2)
```

Ts.boot2 is plotted
```{r, echo = False}
plot(ts.boot2, type = "l")
```
We need to start with a sample of a certain size B of consecutive points from a time series ts.series. First we need to randomly select a single point in the time series, and we can do this by selecting a number between 1 and 100.

```{r, echo = False}
ts.sim <- arima.sim(list(order = c(1, 0, 0), ar = 0.5), n = 100) 
single_point <- function(ts.series)
{
  point1 <- sample(c(1:length(ts.series)), 1, replace = T)
  print(ts.sim)
  print(point1)
}
single_point(ts.sim)
```

one_sample creates the sample that we'll be working with now

```{r, echo = False}
one_sample <- function(ts.series, B)
{
  point1 <- sample(c(1:length(ts.series)), 1, replace = T)
  ts.samp1 <- c()
  for(i in 1 : B)
  {
    ts.samp1 <- c(ts.samp1, ts.sim[point1 + i - 1])
  }
  print(ts.sim)
  plot(ts.sim)
  print(point1)
  print(ts.samp1)
  plot(ts.samp1, type = 'l')
}
one_sample(ts.sim, 10)
```

Now we need to separate the sample into blocks of size b. We are assuming that B is divisible by b.

```{r, echo == False}
blocks <- function(ts.series, B, b)
{
  point1 <- sample(c(1:length(ts.series)), 1, replace = T)
  ts.samp1 <- c()
  for(i in 1 : B)
  {
    ts.samp1 <- c(ts.samp1, ts.sim[point1 + i - 1])
  }
  print(ts.sim)
  plot(ts.sim)
  print(point1)
  print(ts.samp1)
  i <- 1
  ts.block_list <- list()
  while(i < B)
  {
    x <- list(ts.samp1[i : (i + b - 1)])
    ts.block_list <- append(ts.block_list, x)
    i = i + b
  }
  print(ts.block_list)
  plot(ts.samp1, type = 'l')
}
blocks(ts.sim, 20, 4)
```

Randomly samples B/b non-overlapping blocks (with replacement) from ts.block_list, creating a single bootstrapped time series. 
```{r, echo = False}
bootstrap_samp1 <- function(ts.series, B, b)
{
  point1 <- sample(c(1:length(ts.series)), 1, replace = T)
  ts.samp1 <- c()
  for(i in 1 : B)
  {
    ts.samp1 <- c(ts.samp1, ts.sim[point1 + i - 1])
  }
  print(ts.sim)
  plot(ts.sim)
  print(point1)
  print(ts.samp1)
  plot(ts.samp1, type = 'l')
  i <- 1
  ts.block_list <- list()
  while(i < B)
  {
    x <- list(ts.samp1[i : (i + b - 1)])
    ts.block_list <- append(ts.block_list, x)
    i = i + b
  }
  print(ts.block_list)
  ts.bootstrap_1 <- sample(ts.block_list, B/b, replace = T)
  print(ts.bootstrap_1)
  class(ts.boot1)
}
bootstrap_samp1(ts.sim, 20, 4)
```

ts.bootstrap_1 is a list of lists and will be difficult to work with, so we use a nested loop to store the values in ts.bootstrap_2, a simpler numeric vector

```{r, echo = False}
bootstrap_samp2 <- function(ts.series, B, b)
{
  point1 <- sample(c(1:length(ts.series)), 1, replace = T)
  ts.samp1 <- c()
  for(i in 1 : B)
  {
    ts.samp1 <- c(ts.samp1, ts.sim[point1 + i - 1])
  }
  print(ts.sim)
  plot(ts.sim)
  print(point1)
  print(ts.samp1)
  plot(ts.samp1, type = 'l')
  i <- 1
  ts.block_list <- list()
  while(i < B)
  {
    x <- list(ts.samp1[i : (i + b - 1)])
    ts.block_list <- append(ts.block_list, x)
    i = i + b
  }
  print(ts.block_list)
  ts.bootstrap_1 <- sample(ts.block_list, B/b, replace = T)
  print(ts.bootstrap_1)
  class(ts.boot1)
  ts.bootstrap_2 <- c()
  for(i in 1 : (B/b)) {
    for(j in 1 : b) {
      ts.bootstrap_2 <- c(ts.bootstrap_2, ts.bootstrap_1[[i]][[j]])
    }
  }
  print(ts.bootstrap_2)
  class(ts.bootstrap_2)
  plot(ts.bootstrap_2, type = 'l')
}
bootstrap_samp2(ts.sim, 20, 4)
```