### 2022-03-06 Simulation_Study.Rmd
Vary Sample Size with Same Block Length

---
title: "Simulation_Study"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We need all necessary functions:

```{r functions, echo = FALSE}
nmblk <- function(n, l) {
  b <- n / l # assuming that is an integer
  idx <- sample(seq(1, n - l + 1, by = l), b, replace = TRUE) # Sample between 1 and n - l + 1, 
  # separating by l, b times
  c(outer(0:(l - 1), idx, "+")) # Creates a vector listing random blocks of size l with starting points 
  # decided by idx
}

mvblk <- function(n, l) {
  b <- n / l
  idx <- sample(1:(n - l + 1), b, replace = TRUE) # Sample any starting point between 1 and n - 1 + 1
  # they can overlap
  c(outer(0:(l - 1), idx, "+")) # Creates a vector listing random blocks of size l with starting points
  # decided by idx
}

blkboot <- function(tseries, statistic, R, l) {
  n <- length(tseries) 
  ## observed statistic
  stat <- statistic(tseries) #stat stores a certain statistic of tseries using a function
  ## nonmoving block bootstrap
  nm.stat <- replicate(R, statistic(tseries[nmblk(n, l)])) # finds the statistic of nm bootstrap and repeats R times
  ## moving block bootstrap
  mv.stat <- replicate(R, statistic(tseries[mvblk(n, l)])) # finds the statistic of mv bootstrap and repeats R times
  list(stat = stat, nm.stat = nm.stat, mv.stat = mv.stat) # returns list of three values
}

do1rep <- function(n, theta, sd, l, statistic) {
  x <- arima.sim(list(ar = theta), n = n, sd = sd)

  # sd is the standard deviation of the noise -> true_variance(X_t) = sd^2/(1 - theta^2)

  bt <- blkboot(x, statistic, 1000, l)

 # nm_lower_bound with quantile function

  nm_interval = quantile(bt$nm.stat, probs = c(0.025, 0.975))
  mv_interval = quantile(bt$mv.stat, probs = c(0.025, 0.975))

  nm_contains = FALSE
  if (nm_interval[1] < 0 && nm_interval[2] > 0)
  {
    nm_contains = TRUE
  }
  mv_contains = FALSE
  if (mv_interval[1] < 0 && mv_interval[2] > 0)
  {
    mv_contains = TRUE
  }

  c(bt$stat, sd(bt$nm.stat), nm_interval, nm_contains, sd(bt$mv.stat), mv_interval, mv_contains)

    # returns statistics and standard error's for nm and mv respectively, standard error (sd(bt$mv.stat) is estimated as the sqrt(sample_variance(X_t) / n)

}
```

Let's conduct a simulation study with n = 100, ar = 0.5, and sd = 0.1 (var = 0.01)

```{r functions, echo = False}
options(max.print = 1000000)
ar1 <- function(x) cor(x[-1], x[-length(x)])
m1 <- function(x) mean(x)

my_summary <- function(x)
{
  rm = rowMeans(x)
  c(rm, sd(x[1,]), rm[2] - sd(x[1,]), rm[6] - sd(x[1,]), (rm[2] - sd(x[1,])) / sd(x[1,]), (rm[6] - sd(x[1,])) / sd(x[1,]))
}

```

```{r functions, echo = False}
sim <- replicate(100, do1rep(100, 0.5, 0.1, 10, m1))

n_100 <- my_summary(sim)

sim <- replicate(100, do1rep(200, 0.5, 0.1, 10, m1))

n_200 <- my_summary(sim)

sim <- replicate(100, do1rep(300, 0.5, 0.1, 10, m1))

n_300 <- my_summary(sim)

sim <- replicate(100, do1rep(400, 0.5, 0.1, 10, m1))

n_400 <- my_summary(sim)

sim <- replicate(100, do1rep(500, 0.5, 0.1, 10, m1))

n_500 <- my_summary(sim)

sim <- replicate(100, do1rep(600, 0.5, 0.1, 10, m1))

n_600 <- my_summary(sim)

sim <- replicate(100, do1rep(700, 0.5, 0.1, 10, m1))

n_700 <- my_summary(sim)

sim <- replicate(100, do1rep(800, 0.5, 0.1, 10, m1))

n_800 <- my_summary(sim)

sim <- replicate(100, do1rep(900, 0.5, 0.1, 10, m1))

n_900 <- my_summary(sim)

sim <- replicate(100, do1rep(1000, 0.5, 0.1, 10, m1))

n_1000 <- my_summary(sim)

n_constant_l.data <- data.frame(
   n = c(100, 200, 300, 400, 500, 600, 700, 800, 900, 1000),
   ar = c(0.5, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5),
   sigma = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1),
   l = c(10, 10, 10, 10, 10, 10, 10, 10, 10, 10),
   mean = c(n_100[1], n_200[1], n_300[1], n_400[1], n_500[1], n_600[1], n_700[1], n_800[1], n_900[1], n_1000[1]), 
   nm_se = c(n_100[2], n_200[2], n_300[2], n_400[2], n_500[2], n_600[2], n_700[2], n_800[2], n_900[2], n_1000[2]),
   nm_lower = c(n_100[3], n_200[3], n_300[3], n_400[3], n_500[3], n_600[3], n_700[3], n_800[3], n_900[3], n_1000[3]), 
   nm_upper = c(n_100[4], n_200[4], n_300[4], n_400[4], n_500[4], n_600[4], n_700[4], n_800[4], n_900[4], n_1000[4]),
   nm_success = c(n_100[5], n_200[5], n_300[5], n_400[5], n_500[5], n_600[5], n_700[5], n_800[5], n_900[5], n_1000[5]),
   mv_se = c(n_100[6], n_200[6], n_300[6], n_400[6], n_500[6], n_600[6], n_700[6], n_800[6], n_900[6], n_1000[6]),
   mv_lower = c(n_100[7], n_200[7], n_300[7], n_400[7], n_500[7], n_600[7], n_700[7], n_800[7], n_900[7], n_1000[7]), 
   mv_upper = c(n_100[8], n_200[8], n_300[8], n_400[8], n_500[8], n_600[8], n_700[8], n_800[8], n_900[8], n_1000[8]),
   mv_success = c(n_100[9], n_200[9], n_300[9], n_400[9], n_500[9], n_600[9], n_700[9], n_800[9], n_900[9], n_1000[9]),
   true_se = c(n_100[10], n_200[10], n_300[10], n_400[10], n_500[10], n_600[10], n_700[10], n_800[10], n_900[10], n_1000[10]),
   nm_diff = c(n_100[11], n_200[11], n_300[11], n_400[11], n_500[11], n_600[11], n_700[11], n_800[11], n_900[11], n_1000[11]),
   mv_diff = c(n_100[12], n_200[12], n_300[12], n_400[12], n_500[12], n_600[12], n_700[12], n_800[12], n_900[12], n_1000[12]),
   nm_percent_change = c(n_100[13], n_200[13], n_300[13], n_400[13], n_500[13], n_600[13], n_700[13], n_800[13], n_900[13], n_1000[13]),
   mv_percent_change = c(n_100[14], n_200[14], n_300[14], n_400[14], n_500[14], n_600[14], n_700[14], n_800[14], n_900[14], n_1000[14])
)
print(n_constant_l.data)
```