### 2022-04-17 Block-Bootstrap.Rmd
New Tests

---
title: "More Tests" 
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We need all necessary functions:

```{r functions, echo = FALSE}
nmblk <- function(n, l) {
  b <- n / l # assuming that is an integer
  idx <- sample(seq(1, n - l + 1, by = l), b, replace = TRUE) # Sample between 1 and n - l + 1, 
  # separating by l, b times
  c(outer(0:(l - 1), idx, "+")) # Creates a vector listing random blocks of size l with starting points 
  # decided by idx
}

mvblk <- function(n, l) {
  b <- n / l
  idx <- sample(1:(n - l + 1), b, replace = TRUE) # Sample any starting point between 1 and n - 1 + 1
  # they can overlap
  c(outer(0:(l - 1), idx, "+")) # Creates a vector listing random blocks of size l with starting points
  # decided by idx
}

blkboot <- function(tseries, statistic, R, l) {
  n <- length(tseries) 
  ## observed statistic
  stat <- statistic(tseries) #stat stores a certain statistic of tseries using a function
  ## nonmoving block bootstrap
  nm.stat <- replicate(R, statistic(tseries[nmblk(n, l)])) # finds the statistic of nm bootstrap and repeats R times
  ## moving block bootstrap
  mv.stat <- replicate(R, statistic(tseries[mvblk(n, l)])) # finds the statistic of mv bootstrap and repeats R times
  list(stat = stat, nm.stat = nm.stat, mv.stat = mv.stat) # returns list of three values
}

do1rep <- function(n, theta, sd, l, statistic) {
  x <- arima.sim(list(ar = theta), n = n, sd = sd)

  # sd is the standard deviation of the noise -> true_variance(X_t) = sd^2/(1 - theta^2)

  bt <- blkboot(x, statistic, 1000, l)

 # nm_lower_bound with quantile function

  nm_interval = quantile(bt$nm.stat, probs = c(0.025, 0.975))
  mv_interval = quantile(bt$mv.stat, probs = c(0.025, 0.975))

  nm_contains = FALSE
  if (nm_interval[1] < 0 && nm_interval[2] > 0)
  {
    nm_contains = TRUE
  }
  mv_contains = FALSE
  if (mv_interval[1] < 0 && mv_interval[2] > 0)
  {
    mv_contains = TRUE
  }

  c(bt$stat, sd(bt$nm.stat), nm_interval, nm_contains, sd(bt$mv.stat), mv_interval, mv_contains)

    # returns statistics and standard error's for nm and mv respectively, standard error (sd(bt$mv.stat) is estimated as the sqrt(sample_variance(X_t) / n)

}
```

```{r more_functions, echo = False}
options(max.print = 1000000)
ar1 <- function(x) cor(x[-1], x[-length(x)])
m1 <- function(x) mean(x)

my_summary <- function(x)
{
  rm = rowMeans(x)
  c(rm, sd(x[1,]), rm[2] - sd(x[1,]), rm[6] - sd(x[1,]), (rm[2] - sd(x[1,])) / sd(x[1,]), (rm[6] - sd(x[1,])) / sd(x[1,]))
}

```

```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(400, 0.2, 0.1, 20, m1)))
root_0.2_1 <- data.frame(c(400, 20, o))
for(l in 19:16)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.2, 0.1, l, m1)))
  root_0.2_1 <- cbind(root_0.2_1, c(n, l, t))

}

print(root_0.2_1)
```
```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(225, 0.2, 0.1, 15, m1)))
root_0.2_2 <- data.frame(c(225, 15, o))
for(l in 14:11)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.2, 0.1, l, m1)))
  root_0.2_2 <- cbind(root_0.2_2, c(n, l, t))
}

print(root_0.2_2)
```
Observed coverage rates seem to be lower, when n was 169 and 144, coverage rates below .90 were recorded for the non-moving method.

```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(100, 0.2, 0.1, 10, m1)))
root_0.2_3 <- data.frame(c(100, 10, o))
for(l in 9:6)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.2, 0.1, l, m1)))
  root_0.2_3 <- cbind(root_0.2_3, c(n, l, t))
}

print(root_0.2_3)
```
For all n's at or below 100, the coverage rates are below .90.

```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(625, 0.4, 0.1, 25, m1)))
root_0.4_1 <- data.frame(c(625, 25, o))
for(l in 24:21)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, l, m1)))
  root_0.4_1 <- cbind(root_0.4_1, c(n, l, t))

}

print(root_0.4_1)
```

```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(400, 0.4, 0.1, 20, m1)))
root_0.4_2 <- data.frame(c(400, 20, o))
for(l in 19:16)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, l, m1)))
  root_0.4_2 <- cbind(root_0.4_2, c(n, l, t))

}

print(root_0.4_2)
```

```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(225, 0.4, 0.1, 15, m1)))
root_0.4_3 <- data.frame(c(225, 15, o))
for(l in 14:11)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, l, m1)))
  root_0.4_3 <- cbind(root_0.4_3, c(n, l, t))

}

print(root_0.4_3)
```

Observed coverage rates seem to be lower, when n was 144 and 121, both coverage rates recorded were below .90.

```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(100, 0.4, 0.1, 10, m1)))
root_0.4_4 <- data.frame(c(100, 10, o))
for(l in 9:6)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, l, m1)))
  root_0.4_4 <- cbind(root_0.4_4, c(n, l, t))

}

print(root_0.4_4)
```

For all n's at or below 100, the coverage rates are below .90.

```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(900, 0.6, 0.1, 30, m1)))
root_0.6_1 <- data.frame(c(900, 30, o))
for(l in 29:26)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, l, m1)))
  root_0.6_1 <- cbind(root_0.6_1, c(n, l, t))

}

print(root_0.6_1)
```

```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(625, 0.6, 0.1, 25, m1)))
root_0.6_2 <- data.frame(c(625, 25, o))
for(l in 24:21)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, l, m1)))
  root_0.6_2 <- cbind(root_0.6_2, c(n, l, t))

}

print(root_0.6_2)
```
Observed coverage rates seem to be lower, when n was 529, a coverage rate below .90 was recorded for the non-moving method.

```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(400, 0.6, 0.1, 20, m1)))
root_0.6_3 <- data.frame(c(400, 20, o))
for(l in 19:16)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, l, m1)))
  root_0.6_3 <- cbind(root_0.6_3, c(n, l, t))

}

print(root_0.6_3)
```

Observed coverage rates seem to be lower, when n was 324 and 256, coverage rate below .90 was recorded for the non-moving methods.

```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(225, 0.6, 0.1, 15, m1)))
root_0.6_4 <- data.frame(c(225, 15, o))
for(l in 14:11)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, l, m1)))
  root_0.6_4 <- cbind(root_0.6_4, c(n, l, t))

}

print(root_0.6_4)
```
9 out of the 10 coverage rates are below .90.

```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(1600, 0.8, 0.1, 40, m1)))
root_0.8_1 <- data.frame(c(1600, 40, o))
for(l in 39:36)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.8, 0.1, l, m1)))
  root_0.8_1 <- cbind(root_0.8_1, c(n, l, t))

}

print(root_0.8_1)
```

```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(1225, 0.8, 0.1, 35, m1)))
root_0.8_2 <- data.frame(c(1225, 35, o))
for(l in 34:31)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.8, 0.1, l, m1)))
  root_0.8_2 <- cbind(root_0.8_2, c(n, l, t))

}

print(root_0.8_2)
```
```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(900, 0.8, 0.1, 30, m1)))
root_0.8_3 <- data.frame(c(900, 30, o))
for(l in 29:26)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.8, 0.1, l, m1)))
  root_0.8_3 <- cbind(root_0.8_3, c(n, l, t))

}

print(root_0.8_3)
```
Observed coverage rates seem to be lower, when n was 676, coverage rates below .90 was recorded for both methods

```{r functions, echo = False}
o = my_summary(replicate(1000, do1rep(625, 0.8, 0.1, 25, m1)))
root_0.8_4 <- data.frame(c(625, 25, o))
for(l in 24:21)
{
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.8, 0.1, l, m1)))
  root_0.8_4 <- cbind(root_0.8_4, c(n, l, t))

}

print(root_0.8_4)
```
All of the coverage rates are below .90.