### 2022-09-25 Manuscript/Block-Bootstrap.Rmd
Error Bars

---
title: "More Tests" 
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We need all necessary functions:

```{r functions, echo = FALSE}
nmblk <- function(n, l) {
  b <- n / l # assuming that is an integer
  idx <- sample(seq(1, n - l + 1, by = l), b, replace = TRUE) # Sample between 1 and n - l + 1, 
  # separating by l, b times
  c(outer(0:(l - 1), idx, "+")) # Creates a vector listing random blocks of size l with starting points 
  # decided by idx
}

mvblk <- function(n, l) {
  b <- n / l
  idx <- sample(1:(n - l + 1), b, replace = TRUE) # Sample any starting point between 1 and n - 1 + 1
  # they can overlap
  c(outer(0:(l - 1), idx, "+")) # Creates a vector listing random blocks of size l with starting points
  # decided by idx
}

blkboot <- function(tseries, statistic, R, l) {
  n <- length(tseries) 
  ## observed statistic
  stat <- statistic(tseries) #stat stores a certain statistic of tseries using a function
  ## nonmoving block bootstrap
  nm.stat <- replicate(R, statistic(tseries[nmblk(n, l)])) # finds the statistic of nm bootstrap and repeats R times
  ## moving block bootstrap
  mv.stat <- replicate(R, statistic(tseries[mvblk(n, l)])) # finds the statistic of mv bootstrap and repeats R times
  list(stat = stat, nm.stat = nm.stat, mv.stat = mv.stat) # returns list of three values
}

do1rep <- function(n, theta, sd, l, statistic) {
  x <- arima.sim(list(ar = theta), n = n, sd = sd)

  # sd is the standard deviation of the noise -> true_variance(X_t) = sd^2/(1 - theta^2)

  bt <- blkboot(x, statistic, 1000, l)

 # nm_lower_bound with quantile function

  nm_interval = quantile(bt$nm.stat, probs = c(0.025, 0.975))
  mv_interval = quantile(bt$mv.stat, probs = c(0.025, 0.975))

  nm_contains = FALSE
  if (nm_interval[1] < 0 && nm_interval[2] > 0)
  {
    nm_contains = TRUE
  }
  mv_contains = FALSE
  if (mv_interval[1] < 0 && mv_interval[2] > 0)
  {
    mv_contains = TRUE
  }

  c(bt$stat, sd(bt$nm.stat), nm_interval, nm_contains, sd(bt$mv.stat), mv_interval, mv_contains)

    # returns statistics and standard error's for nm and mv respectively, standard error (sd(bt$mv.stat) is estimated as the sqrt(sample_variance(X_t) / n)

}
```

```{r more_functions, echo = False}
options(max.print = 1000000)
ar1 <- function(x) cor(x[-1], x[-length(x)])
m1 <- function(x) mean(x)

my_summary <- function(x)
{
  rm = rowMeans(x)
  c(rm, sd(x[1,]), rm[2] - sd(x[1,]), rm[6] - sd(x[1,]), (rm[2] - sd(x[1,])) / sd(x[1,]), (rm[6] - sd(x[1,])) / sd(x[1,]), rm[5] + qnorm(.975) * sqrt((rm[5] * (1 - rm[5])) / 1000), rm[5] - qnorm(.975) * sqrt((rm[5] * (1 - rm[5])) / 1000),  rm[9] + qnorm(.975) * sqrt((rm[9] * (1 - rm[9])) / 1000), rm[9] - qnorm(.975) * sqrt((rm[9] * (1 - rm[9])) / 1000))
}

```

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(400, 0.2, 0.1, 20, m1)))
root_0.2_1 <- data.frame(c(400, 20, o))
for(l in 19:16)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.2, 0.1, l, m1)))
  root_0.2_1 <- cbind(root_0.2_1, c(n, l, t))

}

print(root_0.2_1)
```

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(225, 0.2, 0.1, 15, m1)))
root_0.2_2 <- data.frame(c(225, 15, o))
for(l in 14:11)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.2, 0.1, l, m1)))
  root_0.2_2 <- cbind(root_0.2_2, c(n, l, t))
}

print(root_0.2_2)
```
Observed coverage rates seem to be lower, all are lower than .92 for n <= 225.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(100, 0.2, 0.1, 10, m1)))
root_0.2_3 <- data.frame(c(100, 10, o))
for(l in 9:6)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.2, 0.1, l, m1)))
  root_0.2_3 <- cbind(root_0.2_3, c(n, l, t))
}

print(root_0.2_3)
```
For all n's at or below 100, the coverage rates are below .90.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(625, 0.4, 0.1, 25, m1)))
root_0.4_1 <- data.frame(c(625, 25, o))
for(l in 24:21)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, l, m1)))
  root_0.4_1 <- cbind(root_0.4_1, c(n, l, t))

}

print(root_0.4_1)
```

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(400, 0.4, 0.1, 20, m1)))
root_0.4_2 <- data.frame(c(400, 20, o))
for(l in 19:16)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, l, m1)))
  root_0.4_2 <- cbind(root_0.4_2, c(n, l, t))

}

print(root_0.4_2)
```
All coverage rates below .92 for n <= 441. Coverage for n = 256 below .90 for both methods.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(225, 0.4, 0.1, 15, m1)))
root_0.4_3 <- data.frame(c(225, 15, o))
for(l in 14:11)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, l, m1)))
  root_0.4_3 <- cbind(root_0.4_3, c(n, l, t))

}

print(root_0.4_3)
```

Observed coverage rates seem to be lower, when n was 144 and 121, both coverage rates recorded were below .90. For n = 225, coverage rate below .90 recorded for moving method.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(100, 0.4, 0.1, 10, m1)))
root_0.4_4 <- data.frame(c(100, 10, o))
for(l in 9:6)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, l, m1)))
  root_0.4_4 <- cbind(root_0.4_4, c(n, l, t))

}

print(root_0.4_4)
```

For all n's at or below 100, the coverage rates are below .90.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(900, 0.6, 0.1, 30, m1)))
root_0.6_1 <- data.frame(c(900, 30, o))
for(l in 29:26)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, l, m1)))
  root_0.6_1 <- cbind(root_0.6_1, c(n, l, t))

}

print(root_0.6_1)
```
Coverage rates seem pretty good at this range of n values.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(625, 0.6, 0.1, 25, m1)))
root_0.6_2 <- data.frame(c(625, 25, o))
for(l in 24:21)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, l, m1)))
  root_0.6_2 <- cbind(root_0.6_2, c(n, l, t))

}

print(root_0.6_2)
```
Coverage rates are still pretty good.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(400, 0.6, 0.1, 20, m1)))
root_0.6_3 <- data.frame(c(400, 20, o))
for(l in 19:16)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, l, m1)))
  root_0.6_3 <- cbind(root_0.6_3, c(n, l, t))

}

print(root_0.6_3)
```

Observed coverage rates seem to be lower, when n was 324, a coverage rate below .90 was recorded for the moving method. When n was 256, both coverage rates were below .90. All coverage rates are below .92.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(225, 0.6, 0.1, 15, m1)))
root_0.6_4 <- data.frame(c(225, 15, o))
for(l in 14:11)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, l, m1)))
  root_0.6_4 <- cbind(root_0.6_4, c(n, l, t))

}

print(root_0.6_4)
```
8 out of the 10 coverage rates are below .90.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(400, 0.2, 0.1, 10, m1)))
constant_0.2_1 <- data.frame(c(400, 10, o))
for(m in 19:16)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.2, 0.1, 10, m1)))
  constant_0.2_1 <- cbind(constant_0.2_1, c(n, 10, t))

}

print(constant_0.2_1)
```
Coverage rates are good.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(300, 0.2, 0.1, 10, m1)))
constant_0.2_2 <- data.frame(c(300, 10, o))
for(m in 14:11)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.2, 0.1, 10, m1)))
  constant_0.2_2 <- cbind(constant_0.2_2, c(n, 10, t))

}

print(constant_0.2_2)
```
Coverage rates are still good.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(200, 0.2, 0.1, 10, m1)))
constant_0.2_3 <- data.frame(c(200, 10, o))
for(m in 9:6)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.2, 0.1, 10, m1)))
  constant_0.2_3 <- cbind(constant_0.2_3, c(n, 10, t))

}

print(constant_0.2_3)
```
For n <= 180, coverage rates are below .92. For n = 140, both coverage rates are below .90, and the moving method has a coverage rate below .90 for n = 120.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(100, 0.2, 0.1, 10, m1)))
constant_0.2_4 <- data.frame(c(100, 10, o))
for(m in 4:1)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.2, 0.1, 10, m1)))
  constant_0.2_4 <- cbind(constant_0.2_4, c(n, 10, t))

}

print(constant_0.2_4)
```
Coverage rates are poor.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(600, 0.4, 0.1, 10, m1)))
constant_0.4_1 <- data.frame(c(600, 10, o))
for(m in 29:26)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, 10, m1)))
  constant_0.4_1 <- cbind(constant_0.4_1, c(n, 10, t))

}

print(constant_0.4_1)
```
Good coverage rates.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(500, 0.4, 0.1, 10, m1)))
constant_0.4_2 <- data.frame(c(500, 10, o))
for(m in 24:21)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, 10, m1)))
  constant_0.4_2 <- cbind(constant_0.4_2, c(n, 10, t))

}

print(constant_0.4_2)
```

Good coverage rates.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(400, 0.4, 0.1, 10, m1)))
constant_0.4_3 <- data.frame(c(400, 10, o))
for(m in 19:16)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, 10, m1)))
  constant_0.4_3 <- cbind(constant_0.4_3, c(n, 10, t))

}

print(constant_0.4_3)
```

Still good.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(300, 0.4, 0.1, 10, m1)))
constant_0.4_4 <- data.frame(c(300, 10, o))
for(m in 14:11)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, 10, m1)))
  constant_0.4_4 <- cbind(constant_0.4_4, c(n, 10, t))

}

print(constant_0.4_4)
```

Coverage rates still pretty good.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(200, 0.4, 0.1, 10, m1)))
constant_0.4_5 <- data.frame(c(200, 10, o))
for(m in 9:6)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, 10, m1)))
  constant_0.4_5 <- cbind(constant_0.4_5, c(n, 10, t))

}

print(constant_0.4_5)
```

Non-moving method resulted in coverage rate below .90 for n = 200. All coverage rates for n = 140 and 120 are below .90.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(100, 0.4, 0.1, 10, m1)))
constant_0.4_6 <- data.frame(c(100, 10, o))
for(m in 4:1)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.4, 0.1, 10, m1)))
  constant_0.4_6 <- cbind(constant_0.4_6, c(n, 10, t))

}

print(constant_0.4_6)
```

Poor coverage rates.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(900, 0.6, 0.1, 10, m1)))
constant_0.6_1 <- data.frame(c(900, 10, o))
for(m in 44:41)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, 10, m1)))
  constant_0.6_1 <- cbind(constant_0.6_1, c(n, 10, t))

}

print(constant_0.6_1)
```

All coverage rates below .92.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(800, 0.6, 0.1, 10, m1)))
constant_0.6_2 <- data.frame(c(800, 10, o))
for(m in 39:36)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, 10, m1)))
  constant_0.6_2 <- cbind(constant_0.6_2, c(n, 10, t))

}

print(constant_0.6_2)
```

Moving coverage rate for n = 760 is below .90.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(700, 0.6, 0.1, 10, m1)))
constant_0.6_3 <- data.frame(c(700, 10, o))
for(m in 34:31)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, 10, m1)))
  constant_0.6_3 <- cbind(constant_0.6_3, c(n, 10, t))

}

print(constant_0.6_3)
```
All coverage rates below .92 except those for n = 680.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(600, 0.6, 0.1, 10, m1)))
constant_0.6_4 <- data.frame(c(600, 10, o))
for(m in 29:26)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, 10, m1)))
  constant_0.6_4 <- cbind(constant_0.6_4, c(n, 10, t))

}

print(constant_0.6_4)
```

All coverage rates below .92 except those for n = 540.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(500, 0.6, 0.1, 10, m1)))
constant_0.6_5 <- data.frame(c(500, 10, o))
for(m in 24:21)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, 10, m1)))
  constant_0.6_5 <- cbind(constant_0.6_5, c(n, 10, t))

}

print(constant_0.6_5)
```

Coverage rates are generally good.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(400, 0.6, 0.1, 10, m1)))
constant_0.6_6 <- data.frame(c(400, 10, o))
for(m in 19:16)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, 10, m1)))
  constant_0.6_6 <- cbind(constant_0.6_6, c(n, 10, t))

}

print(constant_0.6_6)
```

All coverage rates below .92 for n <= 380. Coverage rates below .90 for n = 360 and 340.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(300, 0.6, 0.1, 10, m1)))
constant_0.6_7 <- data.frame(c(300, 10, o))
for(m in 14:11)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, 10, m1)))
  constant_0.6_7 <- cbind(constant_0.6_7, c(n, 10, t))

}

print(constant_0.6_7)
```

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(200, 0.6, 0.1, 10, m1)))
constant_0.6_8 <- data.frame(c(200, 10, o))
for(m in 9:6)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0.6, 0.1, 10, m1)))
  constant_0.6_8 <- cbind(constant_0.6_8, c(n, 10, t))

}

print(constant_0.6_8)
```

```{r} 
colnames(root_0.2_1) <- NULL
colnames(root_0.2_2) <- NULL
colnames(root_0.2_3) <- NULL
root_0.2_3
```

```{r}

root_0.2 <- data.frame(rbind(t(root_0.2_1[c(1, 2, 7, 18, 17),]), t(root_0.2_1[c(1, 2, 11, 20, 19),]), t(root_0.2_2[c(1, 2, 7, 18, 17),]), t(root_0.2_2[c(1, 2, 11, 20, 19),]), t(root_0.2_3[c(1, 2, 7, 18, 17),]), t(root_0.2_3[c(1, 2, 11, 20, 19),])))
rownames(root_0.2) <- NULL
colnames(root_0.2) <- c("n", "l", 'coverage_rate', 'lb', 'ub')
root_0.2['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
root_0.2[order(-root_0.2$n), ]
```
```{r}
library(ggplot2)
ggplot(root_0.2, aes(x = n, y = coverage_rate, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=5) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = ' * sqrt(n) * ', AR = 0.2'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('root_0.2.pdf')
```

```{r}
root_0.4 <- data.frame(rbind(t(root_0.4_1[c(1, 2, 7, 18, 17),]), t(root_0.4_1[c(1, 2, 11, 20, 19),]), t(root_0.4_2[c(1, 2, 7, 18, 17),]), t(root_0.4_2[c(1, 2, 11, 20, 19),]), t(root_0.4_3[c(1, 2, 7, 18, 17),]), t(root_0.4_3[c(1, 2, 11, 20, 19),]), t(root_0.4_4[c(1, 2, 7, 18, 17),]), t(root_0.4_4[c(1, 2, 11, 20, 19),])))
rownames(root_0.4) <- NULL
colnames(root_0.4) <- c("n", "l", 'ci', 'lb', 'ub')
root_0.4['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
root_0.4[order(-root_0.4$n), ]
```
```{r}
library(ggplot2)
ggplot(root_0.4, aes(x = n, y = ci, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=5) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = ' * sqrt(n) * ', AR = 0.4'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('root_0.4.pdf')
```
```{r}
root_0.6 <- data.frame(rbind(t(root_0.6_1[c(1, 2, 7, 18, 17),]), t(root_0.6_1[c(1, 2, 11, 20, 19),]), t(root_0.6_2[c(1, 2, 7, 18, 17),]), t(root_0.6_2[c(1, 2, 11, 20, 19),]), t(root_0.6_3[c(1, 2, 7, 18, 17),]), t(root_0.6_3[c(1, 2, 11, 20, 19),]), t(root_0.6_4[c(1, 2, 7, 18, 17),]), t(root_0.6_4[c(1, 2, 11, 20, 19),])))
rownames(root_0.6) <- NULL
colnames(root_0.6) <- c("n", "l", 'ci', 'lb', 'ub')
root_0.6['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
root_0.6[order(-root_0.6$n), ]
```

```{r}
library(ggplot2)
ggplot(root_0.6, aes(x = n, y = ci, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=5) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = ' * sqrt(n) * ', AR = 0.6'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('root_0.6.pdf')
```

```{r}
constant_0.2 <- data.frame(rbind(t(constant_0.2_1[c(1, 2, 7, 18, 17),]), t(constant_0.2_1[c(1, 2, 11, 20, 19),]), t(constant_0.2_2[c(1, 2, 7, 18, 17),]), t(constant_0.2_2[c(1, 2, 11, 20, 19),]), t(constant_0.2_3[c(1, 2, 7, 18, 17),]), t(constant_0.2_3[c(1, 2, 11, 20, 19),]), t(constant_0.2_4[c(1, 2, 7, 18, 17),]), t(constant_0.2_4[c(1, 2, 11, 20, 19),])))
rownames(constant_0.2) <- NULL
colnames(constant_0.2) <- c("n", "l", 'ci', 'lb', 'ub')
constant_0.2['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
constant_0.2[order(-constant_0.2$n), ]
```

```{r}
library(ggplot2)
ggplot(constant_0.2, aes(x = n, y = ci, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=5) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = 10, AR = 0.2'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('constant_0.2.pdf')
```

```{r}
constant_0.4 <- data.frame(rbind(t(constant_0.4_1[c(1, 2, 7, 18, 17),]), t(constant_0.4_1[c(1, 2, 11, 20, 19),]), t(constant_0.4_2[c(1, 2, 7, 18, 17),]), t(constant_0.4_2[c(1, 2, 11, 20, 19),]), t(constant_0.4_3[c(1, 2, 7, 18, 17),]), t(constant_0.4_3[c(1, 2, 11, 20, 19),]), t(constant_0.4_4[c(1, 2, 7, 18, 17),]), t(constant_0.4_4[c(1, 2, 11, 20, 19),]), t(constant_0.4_5[c(1, 2, 7, 18, 17),]), t(constant_0.4_5[c(1, 2, 11, 20, 19),]), t(constant_0.4_6[c(1, 2, 7, 18, 17),]), t(constant_0.4_6[c(1, 2, 11, 20, 19),])))
rownames(constant_0.4) <- NULL
colnames(constant_0.4) <- c("n", "l", 'ci', 'lb', 'ub')
constant_0.4['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
constant_0.4[order(-constant_0.4$n), ]
```

```{r}
library(ggplot2)
ggplot(constant_0.4, aes(x = n, y = ci, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=5) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = 10, AR = 0.4'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('constant_0.4.pdf')
```

```{r}
constant_0.6 <- data.frame(rbind(t(constant_0.6_1[c(1, 2, 7, 18, 17),]), t(constant_0.6_1[c(1, 2, 11, 20, 19),]), t(constant_0.6_2[c(1, 2, 7, 18, 17),]), t(constant_0.6_2[c(1, 2, 11, 20, 19),]), t(constant_0.6_3[c(1, 2, 7, 18, 17),]), t(constant_0.6_3[c(1, 2, 11, 20, 19),]), t(constant_0.6_4[c(1, 2, 7, 18, 17),]), t(constant_0.6_4[c(1, 2, 11, 20, 19),]), t(constant_0.6_5[c(1, 2, 7, 18, 17),]), t(constant_0.6_5[c(1, 2, 11, 20, 19),]), t(constant_0.6_6[c(1, 2, 7, 18, 17),]), t(constant_0.6_6[c(1, 2, 11, 20, 19),]), t(constant_0.6_7[c(1, 2, 7, 18, 17),]), t(constant_0.6_7[c(1, 2, 11, 20, 19),]), t(constant_0.6_8[c(1, 2, 7, 18, 17),]), t(constant_0.6_8[c(1, 2, 11, 20, 19),])))
rownames(constant_0.6) <- NULL
colnames(constant_0.6) <- c("n", "l", 'ci', 'lb', 'ub')
constant_0.6['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
constant_0.6[order(-constant_0.6$n), ]
```

```{r}
library(ggplot2)
ggplot(constant_0.6, aes(x = n, y = ci, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=5) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = 10, AR = 0.6'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('constant_0.6.pdf')
```