### 2022-11-23 Code/Visualizations.Rmd

---
title: "Visualization"
output: html_document
date: "2022-09-30"
---

```{r}
set.seed(5)
(x <- arima.sim(list(ar = 0.2), n = 50, sd = 1)) 
pdf("orig_ts.pdf", width = 15)
plot.ts(x)
text(x=c(5, 15, 25), y=c(2, 2, 2), labels=c("Block 3, 5", "Block 1", "Block 2, 4"), col="blue")
rect(1, -3, 10, 3, col= rgb(0,0,0,alpha=0.25))
rect(11, -3, 20, 3, col= rgb(0,0,1,alpha=0.25))
rect(21, -3, 30, 3, col= rgb(0,1,0,alpha=0.25))
rect(31, -3, 40, 3, col= rgb(0,1,1,alpha=0))
rect(41, -3, 50, 3, col= rgb(1,0,0,alpha=0))
dev.off()
```

```{r}
set.seed(5)
(y1 <- x[nmblk(50, 10)])
pdf("nm_ts.pdf", width = 15)
plot.ts(y1)
text(x=c(5, 15, 25, 35, 45), y=c(2, 2, 2, 2, 2), labels=c("Block 1", "Block 2", "Block 3", "Block 4", "Block 5"), col="blue")
rect(1, -3, 10, 3, col= rgb(0,0,1,alpha=0.25))
rect(11, -3, 20, 3, col= rgb(0,1,0,alpha=0.25))
rect(21, -3, 30, 3, col= rgb(0,0,0,alpha=0.25))
rect(31, -3, 40, 3, col= rgb(0,1,0,alpha=0.25))
rect(41, -3, 50, 3, col= rgb(0,0,0,alpha=0.25))
dev.off()
```
```{r}
set.seed(5)
(y2 <- x[mvblk(50, 10)])
pdf("mv_ts.pdf", width = 15)
plot.ts(y2)
text(x=c(5, 15, 25, 35, 45), y=c(2, 2, 2, 2, 2), labels=c("Block 1", "Block 2", "Block 3", "Block 4", "Block 5"), col="blue")
rect(1, -3, 10, 3, col= rgb(0,0,0,alpha=0.25))
rect(11, -3, 20, 3, col= rgb(0,0,1,alpha=0.25))
rect(21, -3, 30, 3, col= rgb(0,1,0,alpha=0.25))
rect(31, -3, 40, 3, col= rgb(0,1,1,alpha=0.25))
rect(41, -3, 50, 3, col= rgb(1,0,0,alpha=0.25))
dev.off()
```
```{r}
pdf("orig_ts_2.pdf", width = 15)
plot.ts(x)
rect(2, -3, 11, 3, col= rgb(0,0,0,alpha=0.25))
rect(15, -3, 24, 3, col= rgb(0,0,1,alpha=0.5))
rect(11, -3, 20, 3, col= rgb(0,1,0,alpha=0.25))
rect(41, -3, 50, 3, col= rgb(0,1,1,alpha=0.25))
rect(21, -3, 30, 3, col= rgb(1,0,0,alpha=0.25))
text(x=c(6, 19, 15, 45, 25), y=c(2, -2, 2, 2, 2), labels=c("Block 1", "Block 2", "Block 3", "Block 4", "Block 5"), col="blue")
dev.off()
```
```{r}
set.seed(5)
data <- blkboot(x, m1, 1000, 10)
library(ggplot2)
pdf('nm_dist.pdf')
hist(data$nm.stat, main = 'Distribution of Sample Means from Non-Moving Block Bootstrap', xlab = 'Sample Mean')
dev.off()
pdf('mv_dist.pdf')
hist(data$mv.stat, main = 'Distribution of Sample Means from Moving Block Bootstrap', xlab = 'Sample Mean')
dev.off()
```