### 2022-11-28 Code/Plots.Rmd
Plots

```{r}
(m_0 <- read.csv('../Data/m_0.csv'))
```
```{r}
(m_0.2 <- read.csv('../Data/m_0.2.csv'))
```
```{r}
(m_0.4 <- read.csv('../Data/m_0.4.csv'))
```
```{r}
(ar_0 <- read.csv('../Data/ar_0.csv'))
```
```{r}
(ar_0.2 <- read.csv('../Data/ar_0.2.csv'))
```
```{r}
(ar_0.4 <- read.csv('../Data/ar_0.4.csv'))
```
```{r}
library(ggplot2)
library(gridExtra)
m_0_nm <- ggplot(m_0, aes(x = Sample.Length, y = Non.Moving.Coverage.Rate)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Non.Moving.Lower.Bound, ymax=Non.Moving.Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(title = 'Theta = 0, Non-Moving Method for Mean', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate') +
  theme(legend.position="none") + 
  scale_x_continuous(trans='log2')
m_0_nm
```
```{r}
m_0_m <- ggplot(m_0, aes(x = Sample.Length, y = Moving.Coverage.Rate)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Moving.Lower.Bound, ymax=Moving.Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(title = 'Theta = 0, Moving Method for Mean', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate') +
  theme(legend.position="none") + 
  scale_x_continuous(trans='log2')
grid.arrange(m_0_nm, m_0_m, ncol=2)
```
```{r}
ggplot(m_0.2, aes(x = Sample.Length, y = Non.Moving.Coverage.Rate)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Non.Moving.Lower.Bound, ymax=Non.Moving.Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(title = 'Theta = 0.2, Non-Moving Method for Mean', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate') +
  theme(legend.position="none") +
  scale_x_continuous(trans='log2')
```
```{r}
ggplot(m_0.2, aes(x = Sample.Length, y = Moving.Coverage.Rate)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Moving.Lower.Bound, ymax=Moving.Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(title = 'Theta = 0.2, Moving Method for Mean', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate') +
  theme(legend.position="none") +
  scale_x_continuous(trans='log2')
```
```{r}
ggplot(m_0.4, aes(x = Sample.Length, y = Non.Moving.Coverage.Rate)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Non.Moving.Lower.Bound, ymax=Non.Moving.Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(title = 'Theta = 0.4, Non-Moving Method for Mean', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate') +
  theme(legend.position="none") +
  scale_x_continuous(trans='log2')
```
```{r}
ggplot(ar_0, aes(x = Sample.Length, y = Non.Moving.Coverage.Rate)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Non.Moving.Lower.Bound, ymax=Non.Moving.Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(title = 'Theta = 0, Non-Moving Method for AR(1) Coefficient', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate') +
  theme(legend.position="none") +
  scale_x_continuous(trans='log2')
```
```{r}
ggplot(ar_0, aes(x = Sample.Length, y = Moving.Coverage.Rate)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Moving.Lower.Bound, ymax=Moving.Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(title = 'Theta = 0, Moving Method for AR(1) Coefficient', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate') +
  theme(legend.position="none") +
  scale_x_continuous(trans='log2')
```
```{r}
ggplot(ar_0.2, aes(x = Sample.Length, y = Non.Moving.Coverage.Rate)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Non.Moving.Lower.Bound, ymax=Non.Moving.Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(title = 'Theta = 0.2, Non-Moving Method for AR(1) Coefficient', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate') +
  theme(legend.position="none") +
  scale_x_continuous(trans='log2')
```
```{r}
ggplot(ar_0.2, aes(x = Sample.Length, y = Moving.Coverage.Rate)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Moving.Lower.Bound, ymax=Moving.Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(title = 'Theta = 0.2, Moving Method for AR(1) Coefficient', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate') +
  theme(legend.position="none") +
  scale_x_continuous(trans='log2')
```
```{r}
ggplot(ar_0.4, aes(x = Sample.Length, y = Non.Moving.Coverage.Rate)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Non.Moving.Lower.Bound, ymax=Non.Moving.Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(title = 'Theta = 0.4, Non-Moving Method for AR(1) Coefficient', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate') +
  theme(legend.position="none") +
  scale_x_continuous(trans='log2')
```
```{r}
ggplot(ar_0.4, aes(x = Sample.Length, y = Moving.Coverage.Rate)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Moving.Lower.Bound, ymax=Moving.Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  labs(title = 'Theta = 0.4, Moving Method for AR(1) Coefficient', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate') +
  theme(legend.position="none") +
  scale_x_continuous(trans='log2')
```
```{r}
(df <- read.csv('../Data/df.csv'))
```
```{r}
library(ggplot2)
ggplot(df[df$Theta == 0, ], aes(x = Sample.Length, y = Coverage.Rate, color = Method)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Lower.Bound, ymax=Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  facet_grid(Method ~ Target) +
  labs(title = 'Theta = 0', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_x_continuous(trans='log2')
ggsave('../Manuscript/theta_0.pdf', height = 5)
```
```{r}
library(ggplot2)
ggplot(df[df$Theta == 0.2, ], aes(x = Sample.Length, y = Coverage.Rate, color = Method)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Lower.Bound, ymax=Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  facet_grid(Method ~ Target) +
  labs(title = 'Theta = 0.2', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_x_continuous(trans='log2')
ggsave('../Manuscript/theta_0.2.pdf', height = 5)
```
```{r}
library(ggplot2)
ggplot(df[df$Theta == 0.4, ], aes(x = Sample.Length, y = Coverage.Rate, color = Method)) +
  geom_hline(yintercept = .95, linetype = 'dashed', color = 'orange') + 
  geom_errorbar(aes(ymin=Lower.Bound, ymax=Upper.Bound), colour="black", width=0.1) +
  geom_line() +
  geom_point(size = 0.5) +
  facet_grid(Method ~ Target) +
  labs(title = 'Theta = 0.4', x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_x_continuous(trans='log2')
ggsave('../Manuscript/theta_0.4.pdf', height = 5)
```