### 2023-01-18 Code/simulation.Rmd

```{r functions}
## This note checks the empirical coverage rate of CIs constructed from
## block bootstrap with series generated from an AR process

mystat <- function(x) {
    c(mean(x),
      sd(x),
      cor(x[-1], x[-length(x)]))
      # acf(x, lag.max = 1, plot = FALSE)$acf[2])
}

do1rep <- function(n, theta, statistic, blksize, R = 1000, level = .95) {
    x <- arima.sim(list(ar = theta), n = n) * sqrt(1 - theta^2)
    bts <- boot::tsboot(x, statistic, l = blksize, sim = "fixed", R = R)
    alpha <- 1 - level

    ## Percentile interval
    mbts <- apply(bts$t, 2, mean)
    crit <- 
      apply(t(t(bts$t) - mbts), 2, quantile, prob = c(alpha/2, 1 - alpha/2)) 
    ## alpha/2 and 1 - alpha/2 critical values 
    ## of pseudo-estimate - mean of pseudo-estimates
    crit2 <- 
      apply(t(t(bts$t) - bts$t0), 2, quantile, prob = c(alpha/2, 1 - alpha/2)) 
    ## alpha/2 and 1 - alpha/2 critical values 
    ## of pseudo-estimate - original estimate

    ## interval centered around...
    estCI <- sweep(crit, 2, bts$t0, FUN = '+') # original estimate
    ## LB: bts$t0 + alpha/2 crit, UB: bts$t0 + (1 - alpha/2) crit
    mbtCI <- sweep(crit2, 2, mbts, FUN = '+') # mean of pseudo-estimates
    ## LB: mbts + alpha/2 crit2, UB: mbts + (1 - alpha/2) crit2
    altCI <- sweep(-crit[order(2:1),], 2, bts$t0, FUN = '+') # both
    ## LB: bts$t0 - (1 - alpha/2) crit, UB: bts$t0 - alpha/2 crit and
    ## LB: mbts - (1 - alpha/2) crit2, UB: mbts - alpha/2 crit2

    ## BCa interval
    z0 <- qnorm(colMeans(sweep(bts$t, 2, bts$t0) < 0))
    thetajack <- sapply(1: (n / blksize), 
                        function(i) 
                          statistic(x[ - ((i - 1) * blksize + 1:blksize)]))
    a <- apply(thetajack, 1, e1071::skewness) / 6
    alpha1 <- pnorm(
      z0 + (z0 + qnorm(alpha/2)) / (1 - a * (z0 + qnorm(alpha/2))))
    alpha2 <- pnorm(
      z0 + (z0 + qnorm(1 - alpha/2)) / (1 - a * (z0 + qnorm(1 - alpha/2))))
    crit3 <- sapply(1:length(bts$t0), function(i) 
      quantile(t(t(bts$t[,i]) - mbts[i]), prob = c(alpha1[i], alpha2[i]))) 
    ## alpha1 and alpha2 critical values 
    ## of pseudo-estimate - mean of pseudo-estimates
    crit4 <- sapply(1:length(bts$t0), function(i) 
      quantile(t(t(bts$t[,i]) - bts$t0[i]), prob = c(alpha1[i], alpha2[i])))
    ## alpha1 and alpha2 critical values 
    ## of pseudo-estimate - original estimate

    ## interval centered around...
    bcaCIest <- sweep(crit3, 2, bts$t0, FUN = '+') # original estimate
    ## LB: bts$t0 + alpha1 crit3, UB: bts$t0 + alpha2 crit3
    bcaCImbt <- sweep(crit4, 2, mbts, FUN = '+') # mean of pseudo-estimates
    ## LB: mbts + alpha1 crit4, UB: mbts + alpha2 crit4
    bcaCIalt <- sweep(-crit3[order(2:1),], 2, bts$t0, FUN = '+') # both
    ## LB: bts$t0 - alpha2 crit3, UB: bts$t0 - alpha1 crit3 and
    ## LB: mbts - alpha2 crit4, UB: mbts - alpha1 crit4

    ## return all intervals
    c(estCI, mbtCI, altCI, bcaCIest, bcaCImbt, bcaCIalt)
}

mychk <- function(sim, target) {
    p <- nrow(sim) / length(target) / 2
    target <- rep(target, p)
    ret <- rep(NA, length(target))
    for (i in seq_along(ret)) {
        ii <- (i - 1) * 2
        ret[i] <- mean(sim[ii + 1,] < target[i] & target[i] < sim[ii + 2,])
    }
    ret
}
```

```{r simulation}
nrep <- 1000
```

```{r ma = .1, n = 100}
n <- 100
blksize <- 5
theta <- .1
target <- c(0, 1, theta)
set.seed(1234)
sim_.1_1h <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.1_1h,"../Data/Raw/sim_.1_1h.csv", row.names = FALSE)
mychk(sim_.1_1h, target)
```
```{r ma = .1, n = 500}
n <- 500
blksize <- 8
theta <- .1
target <- c(0, 1, theta)
set.seed(2345)
sim_.1_5h <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.1_5h,"../Data/Raw/sim_.1_5h.csv", row.names = FALSE)
mychk(sim_.1_5h, target)
```
```{r ma = .1, n = 1000}
n <- 1000
blksize <- 10
theta <- .1
target <- c(0, 1, theta)
set.seed(3456)
sim_.1_1k <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.1_1k,"../Data/Raw/sim_.1_1k.csv", row.names = FALSE)
mychk(sim_.1_1k, target)
```
```{r ma = .3, n = 100}
n <- 100
blksize <- 5
theta <- .3
target <- c(0, 1, theta)
set.seed(1234)
sim_.3_1h <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.3_1h,"../Data/Raw/sim_.3_1h.csv", row.names = FALSE)
mychk(sim_.3_1h, target)
```
```{r ma = .3, n = 500}
n <- 500
blksize <- 8
theta <- .3
target <- c(0, 1, theta)
set.seed(2345)
sim_.3_5h <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.3_5h,"../Data/Raw/sim_.3_5h.csv", row.names = FALSE)
mychk(sim_.3_5h, target)
```
```{r ma = .3, n = 1000}
n <- 1000
blksize <- 10
theta <- .3
target <- c(0, 1, theta)
set.seed(3456)
sim_.3_1k <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.3_1k,"../Data/Raw/sim_.3_1k.csv", row.names = FALSE)
mychk(sim_.3_1k, target)
```
```{r ma = .5, n = 100}
n <- 100
blksize <- 5
theta <- .5
target <- c(0, 1, theta)
set.seed(1234)
sim_.5_1h <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.5_1h,"../Data/Raw/sim_.5_1h.csv", row.names = FALSE)
mychk(sim_.5_1h, target)
```
```{r ma = .5, n = 500}
n <- 500
blksize <- 8
theta <- .5
target <- c(0, 1, theta)
set.seed(2345)
sim_.5_5h <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.5_5h,"../Data/Raw/sim_.5_5h.csv", row.names = FALSE)
mychk(sim_.5_5h, target)
```
```{r ma = .5, n = 1000}
n <- 1000
blksize <- 10
theta <- .5
target <- c(0, 1, theta)
set.seed(3456)
sim_.5_1k <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.5_1k,"../Data/Raw/sim_.5_1k.csv", row.names = FALSE)
mychk(sim_.5_1k, target)
```