### 2023-01-19 Code/NewPlots.Rmd
Data Update

```{r}
graph_bts <- function(cov, theta, target, width, nrep = 1000, sizes = c(100, 500, 1000), types = c('estCI', 'mbtCI', 'altCI', 'bcaCIest', 'bcaCImbt', 'bcaCIalt'), parameters = c('mu', 'sigma', 'theta'), level = .95) {
  n <- rep(sizes, each = length(types) * length(parameters))
  CI <- rep(types, times = length(sizes), each = length(parameters))
  t <- rep(parameters, times = length(sizes) * length(types))
  LB <- sapply(cov, function(i) prop.test(i*nrep, nrep)$conf.int[1])
  UB <-sapply(cov, function(i) prop.test(i*nrep, nrep)$conf.int[2])

  df <- data.frame(n, CI, t, cov, LB, UB)
  library(ggplot2)
  ggplot(df, aes(x = n, y = cov)) +
    geom_hline(yintercept = level, linetype = 'dashed', color = 'orange') + 
    geom_line() +
    geom_errorbar(aes(ymin=LB, ymax = UB), colour="black", width = width) +
    facet_grid(t ~ CI)
}
```

```{r}
theta <- 0.1
target <- c(0, 1, theta)
cov <- c(mychk(sim_.1_1h, target), mychk(sim_.1_5h, target), mychk(sim_.1_1k, target))

graph_bts(cov, theta, target, width = 50)
ggsave('../Manuscript/figures/plot_.1.pdf')
```
```{r}
theta <- 0.3
target <- c(0, 1, theta)
cov <- c(mychk(sim_.3_1h, target), mychk(sim_.3_5h, target), mychk(sim_.3_1k, target))

graph_bts(cov, theta, target, width = 25)
ggsave('../Manuscript/figures/plot_.3.pdf')
```
```{r}
theta <- 0.5
target <- c(0, 1, theta)
cov <- c(mychk(sim_.5_1h, target), mychk(sim_.5_5h, target), mychk(sim_.5_1k, target))

graph_bts(cov, theta, target, width = 10)
ggsave('../Manuscript/figures/plot_.5.pdf')
```