### 2023-02-09 Code/simulation.Rmd

```{r functions}
## This note checks the empirical coverage rate of CIs constructed from
## block bootstrap with series generated from an AR process

mystat <- function(x) {
    c(mean(x),
      sd(x),
      cor(x[-1], x[-length(x)]))
      # acf(x, lag.max = 1, plot = FALSE)$acf[2])
}

do1rep <- function(n, phi, statistic, blksize, R = 1000, level = .95) {
    if (phi == 0) {
      x <- rnorm(n)
    }
    else {
      x <- arima.sim(list(ar = phi), n = n) * sqrt(1 - phi^2)
    }
    bts <- boot::tsboot(x, statistic, l = blksize, sim = "fixed", R = R)
    alpha <- 1 - level

    ## Percentile interval
    mbts <- apply(bts$t, 2, mean)
    crit <- 
      apply(t(t(bts$t) - mbts), 2, quantile, prob = c(alpha/2, 1 - alpha/2)) 
    ## alpha/2 and 1 - alpha/2 critical values 
    ## of pseudo-estimate - mean of pseudo-estimates
    # crit2 <- 
    #  apply(t(t(bts$t) - bts$t0), 2, quantile, prob = c(alpha/2, 1 - alpha/2)) 
    ## alpha/2 and 1 - alpha/2 critical values 
    ## of pseudo-estimate - original estimate

    ## interval centered around...
    estCI <- sweep(crit, 2, bts$t0, FUN = '+') # original estimate
    ## LB: bts$t0 + alpha/2 crit, UB: bts$t0 + (1 - alpha/2) crit
    # mbtCI <- sweep(crit2, 2, mbts, FUN = '+') # mean of pseudo-estimates
    ## LB: mbts + alpha/2 crit2, UB: mbts + (1 - alpha/2) crit2
    # altCI <- sweep(-crit[order(2:1),], 2, bts$t0, FUN = '+') # both
    ## LB: bts$t0 - (1 - alpha/2) crit, UB: bts$t0 - alpha/2 crit and
    ## LB: mbts - (1 - alpha/2) crit2, UB: mbts - alpha/2 crit2

    ## BCa interval
    z0 <- qnorm(colMeans(sweep(bts$t, 2, bts$t0) < 0))
    thetajack <- sapply(1: (n / blksize), 
                        function(i) 
                          statistic(x[ - ((i - 1) * blksize + 1:blksize)]))
    a <- apply(thetajack, 1, e1071::skewness) / 6
    alpha1 <- pnorm(
      z0 + (z0 + qnorm(alpha/2)) / (1 - a * (z0 + qnorm(alpha/2))))
    alpha2 <- pnorm(
      z0 + (z0 + qnorm(1 - alpha/2)) / (1 - a * (z0 + qnorm(1 - alpha/2))))
    crit3 <- sapply(1:length(bts$t0), function(i) 
      quantile(t(t(bts$t[,i]) - mbts[i]), prob = c(alpha1[i], alpha2[i]))) 
    ## alpha1 and alpha2 critical values 
    ## of pseudo-estimate - mean of pseudo-estimates
    # crit4 <- sapply(1:length(bts$t0), function(i) 
    #  quantile(t(t(bts$t[,i]) - bts$t0[i]), prob = c(alpha1[i], alpha2[i])))
    ## alpha1 and alpha2 critical values 
    ## of pseudo-estimate - original estimate

    ## interval centered around...
    bcaCIest <- sweep(crit3, 2, bts$t0, FUN = '+') # original estimate
    ## LB: bts$t0 + alpha1 crit3, UB: bts$t0 + alpha2 crit3
    # bcaCImbt <- sweep(crit4, 2, mbts, FUN = '+') # mean of pseudo-estimates
    ## LB: mbts + alpha1 crit4, UB: mbts + alpha2 crit4
    # bcaCIalt <- sweep(-crit3[order(2:1),], 2, bts$t0, FUN = '+') # both
    ## LB: bts$t0 - alpha2 crit3, UB: bts$t0 - alpha1 crit3 and
    ## LB: mbts - alpha2 crit4, UB: mbts - alpha1 crit4

    ## return all intervals
    c(estCI, bcaCIest)
}

mychk <- function(sim, target) {
    p <- nrow(sim) / length(target) / 2
    target <- rep(target, p)
    ret <- rep(NA, length(target))
    for (i in seq_along(ret)) {
        ii <- (i - 1) * 2
        ret[i] <- mean(sim[ii + 1,] < target[i] & target[i] < sim[ii + 2,])
    }
    ret
}
```

```{r simulation}
nrep <- 1000
```

```{r ar = -.4, n = 100}
n <- 100
blksize <- 5
phi <- -.4
target <- c(0, 1, phi)
set.seed(1234)
sim_n.4_1h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.4_1h,"../Data/Raw/sim_n.4_1h.csv", row.names = FALSE)
mychk(sim_n.4_1h, target)
```

```{r ar = -.4, n = 200}
n <- 200
blksize <- 6
phi <- -.4
target <- c(0, 1, phi)
set.seed(2345)
sim_n.4_2h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.4_2h,"../Data/Raw/sim_n.4_2h.csv", row.names = FALSE)
mychk(sim_n.4_2h, target)
```

```{r ar = -.4, n = 300}
n <- 300
blksize <- 7
phi <- -.4
target <- c(0, 1, phi)
set.seed(3456)
sim_n.4_3h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.4_3h,"../Data/Raw/sim_n.4_3h.csv", row.names = FALSE)
mychk(sim_n.4_3h, target)
```

```{r ar = -.4, n = 400}
n <- 400
blksize <- 7
phi <- -.4
target <- c(0, 1, phi)
set.seed(4567)
sim_n.4_4h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.4_4h,"../Data/Raw/sim_n.4_4h.csv", row.names = FALSE)
mychk(sim_n.4_4h, target)
```

```{r ar = -.4, n = 500}
n <- 500
blksize <- 8
phi <- -.4
target <- c(0, 1, phi)
set.seed(5678)
sim_n.4_5h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.4_5h,"../Data/Raw/sim_n.4_5h.csv", row.names = FALSE)
mychk(sim_n.4_5h, target)
```

```{r ar = -.4, n = 600}
n <- 600
blksize <- 9
phi <- -.4
target <- c(0, 1, phi)
set.seed(6789)
sim_n.4_6h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.4_6h,"../Data/Raw/sim_n.4_6h.csv", row.names = FALSE)
mychk(sim_n.4_6h, target)
```

```{r ar = -.4, n = 700}
n <- 700
blksize <- 9
phi <- -.4
target <- c(0, 1, phi)
set.seed(6789)
sim_n.4_7h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.4_7h,"../Data/Raw/sim_n.4_7h.csv", row.names = FALSE)
mychk(sim_n.4_7h, target)
```

```{r ar = -.2, n = 100}
n <- 100
blksize <- 5
phi <- -.2
target <- c(0, 1, phi)
set.seed(1234)
sim_n.2_1h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.2_1h,"../Data/Raw/sim_n.2_1h.csv", row.names = FALSE)
mychk(sim_n.2_1h, target)
```

```{r ar = -.2, n = 200}
n <- 200
blksize <- 6
phi <- -.2
target <- c(0, 1, phi)
set.seed(2345)
sim_n.2_2h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.2_2h,"../Data/Raw/sim_n.2_2h.csv", row.names = FALSE)
mychk(sim_n.2_2h, target)
```

```{r ar = -.2, n = 300}
n <- 300
blksize <- 7
phi <- -.2
target <- c(0, 1, phi)
set.seed(3456)
sim_n.2_3h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.2_3h,"../Data/Raw/sim_n.2_3h.csv", row.names = FALSE)
mychk(sim_n.2_3h, target)
```

```{r ar = -.2, n = 400}
n <- 400
blksize <- 7
phi <- -.2
target <- c(0, 1, phi)
set.seed(4567)
sim_n.2_4h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.2_4h,"../Data/Raw/sim_n.2_4h.csv", row.names = FALSE)
mychk(sim_n.2_4h, target)
```

```{r ar = -.2, n = 500}
n <- 500
blksize <- 8
phi <- -.2
target <- c(0, 1, phi)
set.seed(5678)
sim_n.2_5h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.2_5h,"../Data/Raw/sim_n.2_5h.csv", row.names = FALSE)
mychk(sim_n.2_5h, target)
```

```{r ar = -.2, n = 600}
n <- 600
blksize <- 9
phi <- -.2
target <- c(0, 1, phi)
set.seed(6789)
sim_n.2_6h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.2_6h,"../Data/Raw/sim_n.2_6h.csv", row.names = FALSE)
mychk(sim_n.2_6h, target)
```

```{r ar = -.2, n = 700}
n <- 700
blksize <- 9
phi <- -.2
target <- c(0, 1, phi)
set.seed(6789)
sim_n.2_7h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_n.2_7h,"../Data/Raw/sim_n.2_7h.csv", row.names = FALSE)
mychk(sim_n.2_7h, target)
```

```{r ar = 0, n = 100}
n <- 100
blksize <- 5
phi <- 0
target <- c(0, 1, phi)
set.seed(1234)
sim_0_1h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_0_1h,"../Data/Raw/sim_0_1h.csv", row.names = FALSE)
mychk(sim_0_1h, target)
```

```{r ar = 0, n = 200}
n <- 200
blksize <- 6
phi <- 0
target <- c(0, 1, phi)
set.seed(2345)
sim_0_2h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_0_2h,"../Data/Raw/sim_0_2h.csv", row.names = FALSE)
mychk(sim_0_2h, target)
```

```{r ar = 0, n = 300}
n <- 300
blksize <- 7
phi <- 0
target <- c(0, 1, phi)
set.seed(3456)
sim_0_3h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_0_3h,"../Data/Raw/sim_0_3h.csv", row.names = FALSE)
mychk(sim_0_3h, target)
```

```{r ar = 0, n = 400}
n <- 400
blksize <- 7
phi <- 0
target <- c(0, 1, phi)
set.seed(4567)
sim_0_4h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_0_4h,"../Data/Raw/sim_0_4h.csv", row.names = FALSE)
mychk(sim_0_4h, target)
```

```{r ar = 0, n = 500}
n <- 500
blksize <- 8
phi <- 0
target <- c(0, 1, phi)
set.seed(5678)
sim_0_5h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_0_5h,"../Data/Raw/sim_0_5h.csv", row.names = FALSE)
mychk(sim_0_5h, target)
```

```{r ar = 0, n = 600}
n <- 600
blksize <- 9
phi <- 0
target <- c(0, 1, phi)
set.seed(6789)
sim_0_6h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_0_6h,"../Data/Raw/sim_0_6h.csv", row.names = FALSE)
mychk(sim_0_6h, target)
```

```{r ar = 0, n = 700}
n <- 700
blksize <- 9
phi <- 0
target <- c(0, 1, phi)
set.seed(6789)
sim_0_7h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_0_7h,"../Data/Raw/sim_0_7h.csv", row.names = FALSE)
mychk(sim_0_7h, target)
```

```{r ar = .2, n = 100}
n <- 100
blksize <- 5
phi <- .2
target <- c(0, 1, phi)
set.seed(1234)
sim_.2_1h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.2_1h,"../Data/Raw/sim_.2_1h.csv", row.names = FALSE)
mychk(sim_.2_1h, target)
```

```{r ar = .2, n = 200}
n <- 200
blksize <- 6
phi <- .2
target <- c(0, 1, phi)
set.seed(2345)
sim_.2_2h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.2_2h,"../Data/Raw/sim_.2_2h.csv", row.names = FALSE)
mychk(sim_.2_2h, target)
```

```{r ar = .2, n = 300}
n <- 300
blksize <- 7
phi <- .2
target <- c(0, 1, phi)
set.seed(3456)
sim_.2_3h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.2_3h,"../Data/Raw/sim_.2_3h.csv", row.names = FALSE)
mychk(sim_.2_3h, target)
```

```{r ar = .2, n = 400}
n <- 400
blksize <- 7
phi <- .2
target <- c(0, 1, phi)
set.seed(4567)
sim_.2_4h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.2_4h,"../Data/Raw/sim_.2_4h.csv", row.names = FALSE)
mychk(sim_.2_4h, target)
```

```{r ar = .2, n = 500}
n <- 500
blksize <- 8
phi <- .2
target <- c(0, 1, phi)
set.seed(5678)
sim_.2_5h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.2_5h,"../Data/Raw/sim_.2_5h.csv", row.names = FALSE)
mychk(sim_.2_5h, target)
```

```{r ar = .2, n = 600}
n <- 600
blksize <- 9
phi <- .2
target <- c(0, 1, phi)
set.seed(6789)
sim_.2_6h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.2_6h,"../Data/Raw/sim_.2_6h.csv", row.names = FALSE)
mychk(sim_.2_6h, target)
```

```{r ar = .2, n = 700}
n <- 700
blksize <- 9
phi <- .2
target <- c(0, 1, phi)
set.seed(6789)
sim_.2_7h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.2_7h,"../Data/Raw/sim_.2_7h.csv", row.names = FALSE)
mychk(sim_.2_7h, target)
```

```{r ar = .4, n = 100}
n <- 100
blksize <- 5
phi <- .4
target <- c(0, 1, phi)
set.seed(1234)
sim_.4_1h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.4_1h,"../Data/Raw/sim_.4_1h.csv", row.names = FALSE)
mychk(sim_.4_1h, target)
```

```{r ar = .4, n = 200}
n <- 200
blksize <- 6
phi <- .4
target <- c(0, 1, phi)
set.seed(2345)
sim_.4_2h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.4_2h,"../Data/Raw/sim_.4_2h.csv", row.names = FALSE)
mychk(sim_.4_2h, target)
```

```{r ar = .4, n = 300}
n <- 300
blksize <- 7
phi <- .4
target <- c(0, 1, phi)
set.seed(3456)
sim_.4_3h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.4_3h,"../Data/Raw/sim_.4_3h.csv", row.names = FALSE)
mychk(sim_.4_3h, target)
```

```{r ar = .4, n = 400}
n <- 400
blksize <- 7
phi <- .4
target <- c(0, 1, phi)
set.seed(4567)
sim_.4_4h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.4_4h,"../Data/Raw/sim_.4_4h.csv", row.names = FALSE)
mychk(sim_.4_4h, target)
```

```{r ar = .4, n = 500}
n <- 500
blksize <- 8
phi <- .4
target <- c(0, 1, phi)
set.seed(5678)
sim_.4_5h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.4_5h,"../Data/Raw/sim_.4_5h.csv", row.names = FALSE)
mychk(sim_.4_5h, target)
```

```{r ar = .4, n = 600}
n <- 600
blksize <- 9
phi <- .4
target <- c(0, 1, phi)
set.seed(6789)
sim_.4_6h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.4_6h,"../Data/Raw/sim_.4_6h.csv", row.names = FALSE)
mychk(sim_.4_6h, target)
```

```{r ar = .4, n = 700}
n <- 700
blksize <- 9
phi <- .4
target <- c(0, 1, phi)
set.seed(6789)
sim_.4_7h <- replicate(nrep, do1rep(n, phi, mystat, blksize))
write.csv(sim_.4_7h,"../Data/Raw/sim_.4_7h.csv", row.names = FALSE)
mychk(sim_.4_7h, target)
```